### 2015-04-08 poisson_errker.tex
Bringing everything up to date

\section{Poisson's Equation} 
We seek an approximation of the function $V$ that satisfies the Poisson equation with homogeneous Dirichlet boundary conditions, i.e. 
\begin{equation} \label{eq:PoissonEquation}
	\begin{split}
		\laplacian{V} &= f, \;\; \text{in} \;\; \oucube^s, \\
		V &= 0, \;\; \text{on} \;\; \partial \ucube^s,
	\end{split}
\end{equation}
where $\laplacian$ is the $s$-dimensional Laplace operator and
$\partial \ucube^s$ denotes the boundary of $\ucube^s$,
i.e. $\partial \ucube^s \cup \oucube^s = \ucube^s$.

\subsection{Greens Function}
The Poisson equation is an example of a well-studied partial differential equation. When the domain of interest is the unit cube $\ucube^s$, the solution can be analytically expressed in terms of a Green's function $\mathsf{G}(\vect{x},\vect{y})$ that represents the potential due to a point source placed at the location $\vect{x}$ inside $\ucube^s$.  In other words $\laplacian_{\vect{x}} \mathsf{G}(\vect{x},\vect{y}) = \delta(\vect{x}-\vect{y})$. The Green's function $\mathsf{G}$ has the Fourier Sine series expansion 
\begin{equation}
	\mathsf{G}(\vect{x},\vect{y}) = \sum_{\vect{m} \in \ints^s_+} 
	\frac{\prod_{i=1}^s \sin(m_i \pi x_i) \sin(m_i \pi y_i)}
	{-\pi^2 \norm{\vect{m}}^2},
	\label{eq:green}
\end{equation}
and the solution $V$ is given by
\begin{equation}
	V(\vect{x}) = \int_{\ucube^s}f(\vect{y})\mathsf{G}(\vect{x},\vect{y}) d\vect{y}.
	\label{eq:analytic}
\end{equation}

@@

Suppose we are given an $L_2(\ucube^s)$ function $\rho(x)$ that is
defined on the open unit cube $\oucube^s$. In order to obtain a
Fourier sine series that converges to $\rho$ almost everywhere in
$\oucube^s$, we need to extend $\rho$ periodically such that it is odd
with respect to each variable. Particularly, let us extend the domain
of $\rho$ so that, within the interval $\cube^s$, it satisfies
\begin{equation}
  \small{
 \rho(\vect{x}) :=  
\begin{cases}
\bigl( \prod_{i=1}^s \sgn(x_i) \bigr)
\rho(|x_1|, \ldots, |x_s| ) &\mbox{if } \forall i \; |x_i| \ne 0 \\ 
0 &\mbox{otherwise},
\end{cases}
}
\label{eq:oddextension}
\end{equation}
while outside $\cube^s$, it is $\cube^s$-periodic, i.e. $\rho(\vect{x}
+ 2\vect{k}) = \rho(\vect{x})$ for $\vect{k} \in \ints^s$. The
extended function $\rho$ can be developed into a multidimensional
Fourier sine series, the coefficients of which are given by 
\begin{equation}
  \st{\rho}[\vect{m}] = 2^s \innerproduct{\rho(\vect{x})}
  {\textstyle{\prod_{i=1}^s}\sin(m_i\pi x_i)}, \; \text{for}\; \vect{m} \in \ints^s_+.
\end{equation}
The coefficient sequence $\st{\rho}[\vect{m}]$ can also be seen as a
special case of the more general multidimensional Fourier
series. In fact, with an odd extension of the sequence $\st{\rho}$, the
function $\rho$ can be expressed as a Fourier series. In particular,
if we define the Fourier series coefficients $\ft{\rho}[\vect{m}]$
(for $\vect{m} \in \ints^s$) as
\begin{equation}
  \small{
  \ft{\rho}[\vect{m}] := 
\begin{cases}
\frac{1}{(2\imath)^{s}} \prod_i \sgn(m_i)
\st{\rho}[|m_1|,\ldots,|m_s|] &\mbox{if } \forall i \; |m_i| \ne 0 \\
0 &\mbox{otherwise},
\end{cases}
}
\label{eq:oddFourierExtension}
\end{equation}
then  $\rho$ is also given by the Fourier series
$
  \rho(\vect{x}) = \sum_{\vect{m} \in
    \ints^s} \ft{\rho}[\vect{m}] 
\exp(\imath \pi \vect{m} \cdot \vect{x})
$.
Thus, we use the notation $\st{\rho}[\cdot]$ and $\ft{\rho}[\cdot]$ to
distinguish between the Fourier sine series and Fourier series
coefficients of the periodic function $\rho$.
%Note that in the above equation, $\ft{\rho}[\vect{m}]$
%denotes the Fourier series of $\rho$ and is valid for all $\vect{m} \in
%\ints^s$.

@@

\subsection{Approximate Solution}
We are interested in the scenario where the function $f$ is only known
through its point samples. Specifically, we assume that the samples of
$f$ reside on the sites of the lattice $\lattice{L}_h$. For the
purpose of enumerating the samples that are contained within
$\cube^s$, let us define the point set
\begin{equation}
  \pointset{P}_h := \underbrace{\{\vect{x}_1,\ldots,\vect{x}_{2^sN} \}}_{\pointset{I}_h}
  \cup
\underbrace{\{\vect{x}_{2^sN + 1},\ldots,\vect{x}_{2^sN + M} \}}_{\pointset{B}_h},
\end{equation}
where $\pointset{I}_h := \ocube^s \cap \{\vect{x} + \vect{m} :
\vect{x} \in \lattice{L}_h \cap \oucube^s, \vect{m} \in \ints^s \}$
consists of interior lattice points, while $\pointset{B}_h$ consists
of extended boundary points, i.e. $\pointset{B}_h := (\lattice{L}_h
\cap \partial\cube^s \cap (-1,1]^s) \cup (\lattice{L}_h \cap
\ocube^s\backslash \pointset{I}_h)$. Two dimensional illustrations are
shown in Figure~\ref{fig:pointset}. We note that with the above
definition of $\pointset{I}_h$, the points $\vect{x}_j$ for $j \in
\{1, \ldots, N\}$ are contained in the open unit cube $\oucube^s$
whereas the remaining points ($\vect{x}_j$ for $j \in \{N+1, \ldots,
2^sN\}$) lie outside. Furthermore, we assume
that the lattice $\lattice{L}$ and the sampling rate $h$ are such that
the set $\{\vect{x} + 2\vect{m} : \vect{x} \in \pointset{P}_h,
\vect{m} \in \ints^s\} = \lattice{L}_h$. We remark that this
requirement for $\lattice{L}_h$ is also
satisfied by integration lattices that are commonly used
to devise quadrature rules within the unit cube $\ucube^s$~\cite{nied92}.
\begin{figure}[t]
  \centering
  \subfloat[Cartesian]{\label{fig:cartesian}\input{./figures/2DCartesian}}
  \hspace{2em}
  \subfloat[Quincunx]{\label{fig:quincunx}\input{./figures/2DQuincunx}}
  \caption[Interior and boundary samples]{Illustration of the point set
  $\pointset{P}_h$ on the two
    dimensional Cartesian ($\mat{L} = \bigl( \begin{smallmatrix} 1&0\\
      0&1
    \end{smallmatrix} \bigr)$, $h=\tfrac{1}{8}$) and Quincunx ($\mat{L} = \bigl(
    \begin{smallmatrix} 1&-1\\
      1&1
    \end{smallmatrix} \bigr)$, $h=\tfrac{1}{10}$) lattices. The interior points
    ($\bullet$) belong to the set $\pointset{I}_h$ while the extended
    boundary points ($\circ$) belong to $\pointset{B}_h$.}
  \label{fig:pointset}
\end{figure}

@@

We wish to use the samples of $f$ to seek an approximation $\appf{V}$ of the
function $V$ that solves the Poisson equation~\eqref{eq:PoissonEquation}. Since
$V$ is a periodic function, we follow the recipe of Jacob \textit{et
al.}~\cite{jacob02} and seek an approximation that lies in a space generated by
a periodic reconstruction function. Specifically, we are interested in the case
where $\appf{V}$ lies in the space $\aspace{\lattice{L}_h}{\varphi_p} :=
\mathrm{span}_{\vect{x}_j\in\pointset{P}_h}\{
\varphi_p(\frac{\vect{x}-\vect{x}_j}{h}) \}$ that is spanned by the scaled and
translated versions of a periodic function $\varphi_p$. Our sought-after
approximation is given by
\begin{equation}
  \appf{V}(\vect{x}) = \sum_{\vect{x}_j \in \pointset{P}_h} c[\vect{x}_j] 
\varphi_p\bigl(\frac{\vect{x} - \vect{x}_j}{h}\bigr),
\label{eq:ansatz}
\end{equation}
where $c[\vect{x}_j]$ is an unknown coefficient sequence defined on
the point set $\pointset{P}_h$ that is to be determined from the
samples of $f$. The function
$\varphi_p$ is a periodized version of a generating function $\varphi$
and is defined as
\begin{equation}
\varphi_p(\vect{x}) := \sum_{\vect{m} \in \ints^s} 
\varphi\bigl(\vect{x} - \frac{2}{h} \vect{m} \bigr).
\label{eq:periodization}
\end{equation}

@@

\subsubsection{Error Kernel for Periodic Functions}
\label{sec:periodicErrorKernel}
In many approximation problems involving periodic functions, one is interested
in seeking an approximation $\appf{z}$ of a periodic function $z$ from its
measurements that lie on some sampling lattice. In light of the notations
introduced earlier, suppose that $z \in L_2(\cube^s)$ and is also
$\cube^s$-periodic. Additionally, suppose that the measurements are made at the
locations $\vect{x}_j \in \pointset{P}_h$ so that one can obtain an
approximation $\appf{z}$ that belongs to  $\aspace{\lattice{L}_h}{\varphi_p}$
and --- in a manner similar to the expansion~\eqref{eq:ansatz} --- is given by
$
  \appf{z}(\vect{x}) = \sum_{\vect{x}_j \in \pointset{P}_h}
  \zeta[\vect{x}_j] \varphi_p(\frac{\vect{x}-\vect{x}_j}{h}),
$
where the
coefficient sequence $\zeta[\cdot]$ is obtained through the discrete
measurements
\begin{equation}
  \zeta[\vect{x}_j] = \innerproduct{z(\vect{x})}{\analysis{\varphi}_p(\frac{\vect{x} - \vect{x}_j}{h})}
\label{eq:measurements}
\end{equation}
made with the scaled and shifted versions of a periodic analysis
function $\analysis{\varphi}_p(\vect{x})$ (the periodization operation
is defined by~\eqref{eq:periodization}). The main result of Jacob
\textit{et al.}~\cite{jacob02} states that the mean square
approximation error $\LLnorm{z - \appf{z}}$ at scale $h$ can
be predicted according to
\begin{equation}
  \sqrt{\sum_{\vect{m} \in \ints^s} \norm{\ft{z}[\vect{m}]}^2 E(\frac{h}{2}\vect{m})}
\label{eq:prediction}
\end{equation}
where $\ft{z}[\vect{m}]$ are the Fourier series coefficients of $z$
and $E(\cdot)$ is the error kernel of Blu and Unser~\cite{blu99}
introduced earlier~\eqref{eq:scErrorKernel}.  Remarkably, the same
error kernel can be used to predict the error for periodic and
non-periodic functions alike. The only change that needs to be made is
in the prediction equation. For functions in $L_2(\reel^s)$, the error
is predicted according to the integral in~\eqref{eq:avgError}. On the
other hand, for periodic functions that belong to $L_2(\cube^s)$, the
prediction equation turns into the summation given
in~\eqref{eq:prediction}. We refer the reader to Jacob \textit{et
  al.}~\cite{jacob02} for details.

\subsubsection{Extension to Linear Operators}
We are interested in approximating from the discrete measurements
of $z$, not the function $z$ itself but a function $\Gamma z$ that is obtained
by applying the linear operator $\Gamma$ to $z$. 
% This is exactly the problem we
% are faced with in the case of approximating the analytic
% solution~\eqref{eq:fourierSolution} to Poisson's equation. 
We would like to
extend the error kernel formulation~\eqref{eq:scErrorKernel} so that it will
allow us to quantify the error incurred in approximating $\Gamma z$ from the
measurements of $z$. Towards this end, we assume that the approximation
$\appf{(\Gamma z)}$ lies in a shift-invariant space spanned by the generator
$\varphi_p$, where the coefficients are obtained from the measurements through a
digital filtering operation. Particularly, we seek an approximation that is
given by
\begin{equation}
\appf{(\Gamma z)}(\vect{x}) = C_h \sum_{\vect{x}_j \in \pointset{P}_h} 
(\zeta \pconv \gamma_h)[\vect{x}_j]
\varphi_p(\frac{\vect{x}-\vect{x}_j}{h}),
\label{eq:opApproximation}
\end{equation}
where $\zeta$ denotes the discrete measurements of $z$ obtained
through~\eqref{eq:measurements}, $\gamma$ is a digital filter defined on
the lattice $\lattice{L}$ and $\gamma_h$ is its corresponding scaled
version, i.e. $\gamma_h[\vect{x}_j] := \gamma[\tfrac{\vect{x}_j}{h}]$ where
$\tfrac{\vect{x}_j}{h} \in \lattice{L}$, and $C_h$ is an associated
scaling constant. The digital filter $\gamma$ represents a suitable
discretization of the operator $\Gamma$ on the lattice $\lattice{L}$. It is
to be applied to the measurements through a cyclic convolution
operation (denoted by $\pconv$) on the lattice $\lattice{L}_h$ using a
periodized version of $\gamma_h$. In particular, the convolution operation
in~\eqref{eq:opApproximation} is defined as
\begin{equation}
  (\zeta \pconv \gamma_h)[\vect{x}_j] :=
  \sum_{\vect{x}_k \in \pointset{P}_h}
  \bigl(
    \sum_{\vect{m} \in \ints^s}\gamma[\frac{\vect{x}_k + 2\vect{m}}{h}]
  \bigr)
  \zeta[\vect{x}_j - \vect{x}_k],
\label{eq:cyclicConvolution}
\end{equation}
where $\vect{x}_j \in \lattice{L}_h$. In the above definition,
$\zeta[\cdot]$ is assumed to be $\cube^s$-periodic.  Furthermore, the
resulting sequence $(\zeta \pconv \gamma_h)[\cdot]$ is also
$\cube^s$-periodic.

>>>>

\section{Poisson's Equation} 
Before detailing how to specifically obtain an indicator function from the divergence of $\vec{V}(\vect{x})$, we discuss a general framework for solving Poisson's equation on a rectangular domain. We provide an analysis in $s$-dimensions, as this generalization comes at little cost. Explicitly, we seek an approximation of the function $\chi(\vect{x})$ that satisfies the Poisson equation with homogeneous Dirichlet boundary conditions, i.e. 
\begin{equation} \label{eq:PoissonEquation}
	\begin{split}
		\laplacian{\chi} &= f, \;\; \text{in} \;\; \oucube^s, \\
		\chi &= 0, \;\; \text{on} \;\; \partial \ucube^s,
	\end{split}
\end{equation}
where $\laplacian$ is the $s$-dimensional Laplace operator and
$\partial \ucube^s$ denotes the boundary of $\ucube^s$,
i.e. $\partial \ucube^s \cup \oucube^s = \ucube^s$.

\subsection{Greens Function}
The Poisson equation is an example of a well-studied partial differential equation. When the domain of interest is the unit cube $\ucube^s$, the solution can be analytically expressed in terms of a Green's function $\mathsf{G}(\vect{x},\vect{y})$ that represents the potential due to a point source placed at the location $\vect{x}$ inside $\ucube^s$.  In other words $\laplacian_{\vect{x}} \mathsf{G}(\vect{x},\vect{y}) = \delta(\vect{x}-\vect{y})$. The Green's function $\mathsf{G}$ has the Fourier Sine series expansion 
\begin{equation}
	\mathsf{G}(\vect{x},\vect{y}) = \sum_{\vect{m} \in \ints^s_+} 
	\frac{\prod_{i=1}^s \sin(m_i \pi x_i) \sin(m_i \pi y_i)}
	{-\pi^2 \norm{\vect{m}}^2},
	\label{eq:green}
\end{equation}
and the solution $\chi$ is given by
\begin{equation}
	\chi(\vect{x}) = \int_{\ucube^s}f(\vect{y})\mathsf{G}(\vect{x},\vect{y}) d\vect{y}.
	\label{eq:analytic}
\end{equation}

@@

Suppose we are given an $L_2(\ucube^s)$ function $\rho(
\vect{x})$ that is
defined on the open unit cube $\oucube^s$. In order to obtain a
Fourier sine series that converges to $\rho$ almost everywhere in
$\oucube^s$, we need to extend $\rho$ periodically such that it is odd
with respect to each variable. Particularly, let us extend the domain
of $\rho$ so that, within the interval $\cube^s$, it satisfies
\begin{equation}
  \small{
 \rho(\vect{x}) :=  
\begin{cases}
\bigl( \prod_{i=1}^s \sgn(x_i) \bigr)
\rho(|x_1|, \ldots, |x_s| ) &\mbox{if } \forall i \; |x_i| \ne 0 \\ 
0 &\mbox{otherwise},
\end{cases}
}
\label{eq:oddextension}
\end{equation}
while outside $\cube^s$, it is $\cube^s$-periodic, i.e. $\rho(\vect{x}
+ 2\vect{k}) = \rho(\vect{x})$ for $\vect{k} \in \ints^s$. The
extended function $\rho$ can be developed into a multidimensional
Fourier sine series, the coefficients of which are given by 
\begin{equation}
  \st{\rho}[\vect{m}] = 2^s \innerproduct{\rho(\vect{x})}
  {\textstyle{\prod_{i=1}^s}\sin(m_i\pi x_i)}, \; \text{for}\; \vect{m} \in \ints^s_+.
\end{equation}
The coefficient sequence $\st{\rho}[\vect{m}]$ can also be seen as a
special case of the more general multidimensional Fourier
series. In fact, with an odd extension of the sequence $\st{\rho}$, the
function $\rho$ can be expressed as a Fourier series. In particular,
if we define the Fourier series coefficients $\ft{\rho}[\vect{m}]$
(for $\vect{m} \in \ints^s$) as
\begin{equation}
  \small{
  \ft{\rho}[\vect{m}] := 
\begin{cases}
\frac{1}{(2\imath)^{s}} \prod_i \sgn(m_i)
\st{\rho}[|m_1|,\ldots,|m_s|] &\mbox{if } \forall i \; |m_i| \ne 0 \\
0 &\mbox{otherwise},
\end{cases}
}
\label{eq:oddFourierExtension}
\end{equation}
then  $\rho$ is also given by the Fourier series
$
  \rho(\vect{x}) = \sum_{\vect{m} \in
    \ints^s} \ft{\rho}[\vect{m}] 
\exp(\imath \pi \vect{m} \cdot \vect{x})
$.
Thus, we use the notation $\st{\rho}[\cdot]$ and $\ft{\rho}[\cdot]$ to
distinguish between the Fourier sine series and Fourier series
coefficients of the periodic function $\rho$.
%Note that in the above equation, $\ft{\rho}[\vect{m}]$
%denotes the Fourier series of $\rho$ and is valid for all $\vect{m} \in
%\ints^s$.

@@

\subsection{Approximate Solution}
We are interested in the scenario where the function $f$ is only known
through its point samples. Specifically, we assume that the samples of
$f$ reside on the sites of the lattice $\lattice{L}_h$. For the
purpose of enumerating the samples that are contained within
$\cube^s$, let us define the point set
\begin{equation}
  \pointset{T}_h := \underbrace{\{\vect{x}_1,\ldots,\vect{x}_{2^sN} \}}_{\pointset{I}_h}
  \cup
\underbrace{\{\vect{x}_{2^sN + 1},\ldots,\vect{x}_{2^sN + M} \}}_{\pointset{B}_h},
\end{equation}
where $\pointset{I}_h := \ocube^s \cap \{\vect{x} + \vect{m} :
\vect{x} \in \lattice{L}_h \cap \oucube^s, \vect{m} \in \ints^s \}$
consists of interior lattice points, while $\pointset{B}_h$ consists
of extended boundary points, i.e. $\pointset{B}_h := (\lattice{L}_h
\cap \partial\cube^s \cap (-1,1]^s) \cup (\lattice{L}_h \cap
\ocube^s\backslash \pointset{I}_h)$. Two dimensional illustrations are
shown in Figure~\ref{fig:pointset}. We note that with the above
definition of $\pointset{I}_h$, the points $\vect{x}_j$ for $j \in
\{1, \ldots, N\}$ are contained in the open unit cube $\oucube^s$
whereas the remaining points ($\vect{x}_j$ for $j \in \{N+1, \ldots,
2^sN\}$) lie outside. Furthermore, we assume
that the lattice $\lattice{L}$ and the sampling rate $h$ are such that
the set $\{\vect{x} + 2\vect{m} : \vect{x} \in \pointset{T}_h,
\vect{m} \in \ints^s\} = \lattice{L}_h$. We remark that this
requirement for $\lattice{L}_h$ is also
satisfied by integration lattices that are commonly used
to devise quadrature rules within the unit cube $\ucube^s$~\cite{nied92}.
\begin{figure}[t]
  \centering
  \subfloat[Cartesian]{\label{fig:cartesian}\input{./figures/2DCartesian}}
  \hspace{2em}
  \subfloat[Quincunx]{\label{fig:quincunx}\input{./figures/2DQuincunx}}
  \caption[Interior and boundary samples]{Illustration of the point set
  $\pointset{T}_h$ on the two
    dimensional Cartesian ($\mat{L} = \bigl( \begin{smallmatrix} 1&0\\
      0&1
    \end{smallmatrix} \bigr)$, $h=\tfrac{1}{8}$) and Quincunx ($\mat{L} = \bigl(
    \begin{smallmatrix} 1&-1\\
      1&1
    \end{smallmatrix} \bigr)$, $h=\tfrac{1}{10}$) lattices. The interior points
    ($\bullet$) belong to the set $\pointset{I}_h$ while the extended
    boundary points ($\circ$) belong to $\pointset{B}_h$.}
  \label{fig:pointset}
\end{figure}

@@

We wish to use the samples of $f$ to seek an approximation $\appf{V}$ of the
function $V$ that solves the Poisson equation~\eqref{eq:PoissonEquation}. Since
$V$ is a periodic function, we follow the recipe of Jacob \textit{et
al.}~\cite{jacob02} and seek an approximation that lies in a space generated by
a periodic reconstruction function. Specifically, we are interested in the case
where $\appf{V}$ lies in the space $\aspace{\lattice{L}_h}{\varphi_p} :=
\mathrm{span}_{\vect{x}_j\in\pointset{T}_h}\{
\varphi_p(\frac{\vect{x}-\vect{x}_j}{h}) \}$ that is spanned by the scaled and
translated versions of a periodic function $\varphi_p$. Our sought-after
approximation is given by
\begin{equation}
  \appf{V}(\vect{x}) = \sum_{\vect{x}_j \in \pointset{T}_h} c[\vect{x}_j] 
\varphi_p\bigl(\frac{\vect{x} - \vect{x}_j}{h}\bigr),
\label{eq:ansatz}
\end{equation}
where $c[\vect{x}_j]$ is an unknown coefficient sequence defined on
the point set $\pointset{T}_h$ that is to be determined from the
samples of $f$. The function
$\varphi_p$ is a periodized version of a generating function $\varphi$
and is defined as
\begin{equation}
\varphi_p(\vect{x}) := \sum_{\vect{m} \in \ints^s} 
\varphi\bigl(\vect{x} - \frac{2}{h} \vect{m} \bigr).
\label{eq:periodization}
\end{equation}

@@

\subsubsection{Error Kernel for Periodic Functions}
\label{sec:periodicErrorKernel}
In many approximation problems involving periodic functions, one is interested
in seeking an approximation $\appf{z}$ of a periodic function $z$ from its
measurements that lie on some sampling lattice. In light of the notations
introduced earlier, suppose that $z \in L_2(\cube^s)$ and is also
$\cube^s$-periodic. Additionally, suppose that the measurements are made at the
locations $\vect{x}_j \in \pointset{T}_h$ so that one can obtain an
approximation $\appf{z}$ that belongs to  $\aspace{\lattice{L}_h}{\varphi_p}$
and --- in a manner similar to the expansion~\eqref{eq:ansatz} --- is given by
$
  \appf{z}(\vect{x}) = \sum_{\vect{x}_j \in \pointset{T}_h}
  \zeta[\vect{x}_j] \varphi_p(\frac{\vect{x}-\vect{x}_j}{h}),
$
where the
coefficient sequence $\zeta[\cdot]$ is obtained through the discrete
measurements
\begin{equation}
  \zeta[\vect{x}_j] = \innerproduct{z(\vect{x})}{\analysis{\varphi}_p(\frac{\vect{x} - \vect{x}_j}{h})}
\label{eq:measurements}
\end{equation}
made with the scaled and shifted versions of a periodic analysis
function $\analysis{\varphi}_p(\vect{x})$ (the periodization operation
is defined by~\eqref{eq:periodization}). The main result of Jacob
\textit{et al.}~\cite{jacob02} states that the mean square
approximation error $\LLnorm{z - \appf{z}}$ at scale $h$ can
be predicted according to
\begin{equation}
  \sqrt{\sum_{\vect{m} \in \ints^s} \norm{\ft{z}[\vect{m}]}^2 E(\frac{h}{2}\vect{m})}
\label{eq:prediction}
\end{equation}
where $\ft{z}[\vect{m}]$ are the Fourier series coefficients of $z$
and $E(\cdot)$ is the error kernel of Blu and Unser~\cite{blu99}
introduced earlier~\eqref{eq:scErrorKernel}.  Remarkably, the same
error kernel can be used to predict the error for periodic and
non-periodic functions alike. The only change that needs to be made is
in the prediction equation. For functions in $L_2(\reel^s)$, the error
is predicted according to the integral in~\eqref{eq:avgError}. On the
other hand, for periodic functions that belong to $L_2(\cube^s)$, the
prediction equation turns into the summation given
in~\eqref{eq:prediction}. We refer the reader to Jacob \textit{et
  al.}~\cite{jacob02} for details.

\subsubsection{Extension to Linear Operators}
We are interested in approximating from the discrete measurements
of $z$, not the function $z$ itself but a function $\Gamma z$ that is obtained
by applying the linear operator $\Gamma$ to $z$. 
% This is exactly the problem we
% are faced with in the case of approximating the analytic
% solution~\eqref{eq:fourierSolution} to Poisson's equation. 
We would like to
extend the error kernel formulation~\eqref{eq:scErrorKernel} so that it will
allow us to quantify the error incurred in approximating $\Gamma z$ from the
measurements of $z$. Towards this end, we assume that the approximation
$\appf{(\Gamma z)}$ lies in a shift-invariant space spanned by the generator
$\varphi_p$, where the coefficients are obtained from the measurements through a
digital filtering operation. Particularly, we seek an approximation that is
given by
\begin{equation}
\appf{(\Gamma z)}(\vect{x}) = C_h \sum_{\vect{x}_j \in \pointset{T}_h} 
(\zeta \pconv \gamma_h)[\vect{x}_j]
\varphi_p(\frac{\vect{x}-\vect{x}_j}{h}),
\label{eq:opApproximation}
\end{equation}
where $\zeta$ denotes the discrete measurements of $z$ obtained
through~\eqref{eq:measurements}, $\gamma$ is a digital filter defined on
the lattice $\lattice{L}$ and $\gamma_h$ is its corresponding scaled
version, i.e. $\gamma_h[\vect{x}_j] := \gamma[\tfrac{\vect{x}_j}{h}]$ where
$\tfrac{\vect{x}_j}{h} \in \lattice{L}$, and $C_h$ is an associated
scaling constant. The digital filter $\gamma$ represents a suitable
discretization of the operator $\Gamma$ on the lattice $\lattice{L}$. It is
to be applied to the measurements through a cyclic convolution
operation (denoted by $\pconv$) on the lattice $\lattice{L}_h$ using a
periodized version of $\gamma_h$. In particular, the convolution operation
in~\eqref{eq:opApproximation} is defined as
\begin{equation}
  (\zeta \pconv \gamma_h)[\vect{x}_j] :=
  \sum_{\vect{x}_k \in \pointset{T}_h}
  \bigl(
    \sum_{\vect{m} \in \ints^s}\gamma[\frac{\vect{x}_k + 2\vect{m}}{h}]
  \bigr)
  \zeta[\vect{x}_j - \vect{x}_k],
\label{eq:cyclicConvolution}
\end{equation}
where $\vect{x}_j \in \lattice{L}_h$. In the above definition,
$\zeta[\cdot]$ is assumed to be $\cube^s$-periodic.  Furthermore, the
resulting sequence $(\zeta \pconv \gamma_h)[\cdot]$ is also
$\cube^s$-periodic.
